Forecasting the old-age dependency ratio to determine a sustainable pension age

### 2021-03-15 Reviews/Response_to_reviewers.Rmd

> Thanks. Now fixed.

2. Last line of Page 5 and first line of Page 6. What does the following sentence mean
in its context: “The deaths are estimated using the standard life table approach of
population projection.”

> We have added a reference to @Preston2000 who discuss this method.

3. Page 6, the numerator and denominator of equation (2) should be flipped?

4. Page 8, after equation (9), it is mentioned that “the HU method is applied to these
two new variables.” In the previous description of the HU method, smoothed version
of $y_t(x)$ is modelled in the form of equation (6), whereas here the non-smoothed
$\log[p_t(x)]$ and $\log[r_t(x)]$ are modelled in the form of equation (6). Please clarify.

> No, we only apply the HU method to smoothed functions: $p_t(x)$ and $r_t(x)$ are functions of $s_t^M(x)$ and $s_t^F(x)$ which are both smooth.

5. Page 9. It is mentioned that “The future deaths and births in this equation are
assumed to follow a Poisson distribution, with parameters as a function of future
mortality and fertility rates.” In the observed data, is the Poisson variation in the
deaths and births modelled?

> Yes, the Poisson variation is used in the weighted smoothing. This information has now been added to end of the first paragraph in Section 2.5.

6. Page 14. In the second step of computing the pension age scheme $P$, what does “or
$a_{T+1} - a_T = 1$” mean? Why is $h$ not involved?

> Thanks you for spotting this error. It has now been corrected to $a_{T+h} - a_{T+h-1}=12$ months. It means that the pension age should not jump more than 12 months between years, to prevent pension-age people losing their pension. We have also added this explanation to the paper.

# Reviewer 2 {-}

@@

Second, what does “sustainable” mean here? In Australia, the age pension is financed by general taxes. Therefore, growth of GDP, which finances the pension, seems important. In fact, the percentage of total pension payment in GDP may be a more meaningful policy target to examine than the percentage of elderly in total population for policy makers. The authors may want to give some thoughts here.

Third, do labor force participation and technological growth matter at all? In most rich countries, the long-term growth rate of per capita GDP is lower over time. Hence even if the pension age is raised, the burden of pension payment, which comes from general taxes, may still become heavier if GDP does not grow as fast in the future. A decline in labor force participation rate may have similar effects. The authors may want to consider these extra factors, though they have stated that the approach of this paper is to “illustrate how population forecasting can be used to explore the pension age issue in a simple context” (p.3), and “whether this OADR is the most appropriate level for Australia remains to be examined and is not the focus of this paper” (p.2).

> Thank you for these helpful comments. We had previously stated this very briefly in the second last paragraph of the conclusion. We have now extended that to incorporate some of the points you have made.

> AIHW (2019) states that the proportion of the aged population receiving the pension has remained relatively stable from 2001 to 2018, so we think it is likely that Power (2014) and the private superannuation firms are counting the numbers differently. So we have now removed the citation to Power (2014) and just refer to AIHW (2019).

> The GDP is much more difficult to forecast than age structure, so it is problematic to use it as a target.

### 2021-03-15 PensionAge.Rmd
Added a Makefile and caching
Revision submitted

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, messages = FALSE, warning = FALSE, cache = FALSE)

@@

Some may argue that using OADR as a measure of a financial burden for the next generation could be biased, with a large number of retirees able to rely on self-financed retirement income, such as superannuation. However, 80% of Australians over pension age rely fully or partly on the age pension [@Power14]. Therefore, the proportion of people over working age is still a good representation of the financial burden on future working generations. Nevertheless, using OADR as a single measurement of financial burden may overlook some important economic factors, and @HY12 have suggested the use of an adjusted OADR that takes account of some of these factors in analyzing pension finance. We do not consider these here. Our approach is intended to illustrate how population forecasting can be used to explore the pension age issue in a simple context, and it could be extended to take into account other variables if required.

@@

The ageing trend in Australia, along with retirees' dependence on the age pension, means the cost of the age pension is currently around at 2.7% of Gross Domestic Profit. With an ageing population, the pension age will need to increase or the fiscal burden will become unsustainable.

@@

Following the general procedure proposed in @HB08, we model each of the five components of population change: female mortality rates, male mortality rates, female net migration, male net migration and fertility rates, and simulate these five components into the future. We use the HU method applied to log-fertility, and the product-ratio method applied to log-mortality and net-migration.

@@

  1. Set $a_{T+h}=a_{T+h-1}$.
  2. Increment $a_{T+h}$ by one-month intervals until either $\widehat{O}_{T+h|T} > O^*$ or $a_{T+1} - a_{T}=$ 1 month, where $\widehat{O}_{T+h|T}$ denotes the mean of the simulated $O_{T+h|T}$ values.

We repeat these steps for $h=2,\dots,H$.

@@

This research addresses the pension age issue from a purely demographic perspective. Further analysis should include an economic perspective as well including the impact of the superannuation guarantee and other economic measures to offset the fiscal burden of the age pension.

Finally, we have not allowed for the impact of COVID-19 in this analysis as it is too early to properly estimate its impact on mortality rates and future age structures. It is known that the mortality rates from COVID-19 are greater amongst older people, so it should lead to a decrease in the OADR. However, the relatively small number of deaths in Australia mean that the effect of the pandemic will probably not substantially change our recommendations.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, messages = FALSE, warning = FALSE, cache = TRUE)

@@

Some may argue that using OADR as a measure of a financial burden for the next generation could be biased, with a large number of retirees able to rely on self-financed retirement income, such as superannuation. However, approximately two-thirds of Australians over pension age rely fully or partly on the age pension [@AIHW19], and this figure has remained relatively constant over the last 20 years. Therefore, the proportion of people over working age is still a good representation of the financial burden on future working generations. Nevertheless, using OADR as a single measurement of financial burden may overlook some important economic factors, and @HY12 have suggested the use of an adjusted OADR that takes account of some of these factors in analyzing pension finance. We do not consider these here. Our approach is intended to illustrate how population forecasting can be used to explore the pension age issue in a simple context, and it could be extended to take into account other variables if required.

@@

The ageing trend in Australia, along with retirees' dependence on the age pension, means the cost of the age pension is currently around at 2.7% of Gross Domestic Product. With an ageing population, the pension age will need to increase or the fiscal burden will become unsustainable.

@@

Following the general procedure proposed in @HB08, we model each of the five components of population change: female mortality rates, male mortality rates, female net migration, male net migration and fertility rates, and simulate these five components into the future. We use the HU method applied to log-fertility, and the product-ratio method applied to log-mortality and net-migration. We used weighted penalized splines [@HU07] when estimating the log-fertility and log-mortality curves, where the weights account for Poisson variation in deaths and births. We use loess when estimating the net-migration curves.

@@

  1. Set $a_{T+h}=a_{T+h-1}$.
  2. Increment $a_{T+h}$ by one-month intervals until either $\widehat{O}_{T+h|T} \le O^*$ or $a_{T+h} - a_{T+h-1}=12$ months, where $\widehat{O}_{T+h|T}$ denotes the mean of the simulated $O_{T+h|T}$ values.

We repeat these steps for $h=2,\dots,H$. The second limit in Step 2 arises because the pension age should not jump more than 12 months between years, to prevent pension-age people losing their pension.

@@

This research addresses the pension age issue from a purely demographic perspective. Further analysis should include an economic perspective as well including the impact of the superannuation guarantee and other economic measures to offset the fiscal burden of the age pension. There are several issues to be considered here. First, not all people aged above the pension age will receive the pension, and the ratio of those receiving the pension may decrease as more retiring people can access superannuation.  Second, labour force participation is probably related to age as self-funded retirees may opt out of paid employment before they reach pension age. Hence, raising the pension age may not have as large an effect on OADR as we have estimated. Third, we have not considered the effect of changing per capita GDP which helps fund the pension. If per capita GDP declines, the fiscal burden of the pension may increase even with a stable OADR.

We have also not allowed for the impact of COVID-19 in this analysis as it is too early to properly estimate its impact on mortality rates and future age structures. It is known that the mortality rates from COVID-19 are greater amongst older people, so it should lead to a decrease in the OADR. However, the relatively small number of deaths in Australia mean that the effect of the pandemic will probably not substantially change our recommendations.

### 2021-03-15 PensionAge_anzjs.tex
Revision submitted

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  doublespace]{anzsauth}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
  \usepackage{amssymb}
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Forecasting the old-age dependency ratio to determine a sustainable pension age},
  pdfauthor={Rob J Hyndman\^{}\{1*\}, Yijun Zeng\^{}2, Han Lin Shang\^{}3},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{longtable,booktabs}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
%% Any special functions or other packages can be loaded here.
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

@@

\usepackage{mathptmx}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{anzsj}

@@

Some may argue that using OADR as a measure of a financial burden for the next generation could be biased, with a large number of retirees able to rely on self-financed retirement income, such as superannuation. However, 80\% of Australians over pension age rely fully or partly on the age pension \citep{Power14}. Therefore, the proportion of people over working age is still a good representation of the financial burden on future working generations. Nevertheless, using OADR as a single measurement of financial burden may overlook some important economic factors, and \citet{HY12} have suggested the use of an adjusted OADR that takes account of some of these factors in analyzing pension finance. We do not consider these here. Our approach is intended to illustrate how population forecasting can be used to explore the pension age issue in a simple context, and it could be extended to take into account other variables if required.

@@

The ageing trend in Australia, along with retirees' dependence on the age pension, means the cost of the age pension is currently around at 2.7\% of Gross Domestic Profit. With an ageing population, the pension age will need to increase or the fiscal burden will become unsustainable.

@@

Following the general procedure proposed in \citet{HB08}, we model each of the five components of population change: female mortality rates, male mortality rates, female net migration, male net migration and fertility rates, and simulate these five components into the future. We use the HU method applied to log-fertility, and the product-ratio method applied to log-mortality and net-migration.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Set \(a_{T+h}=a_{T+h-1}\).
\item
  Increment \(a_{T+h}\) by one-month intervals until either \(\widehat{O}_{T+h|T} > O^*\) or \(a_{T+1} - a_{T}=\) 1 month, where \(\widehat{O}_{T+h|T}\) denotes the mean of the simulated \(O_{T+h|T}\) values.
\end{enumerate}

We repeat these steps for \(h=2,\dots,H\).

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  doublespace]{anzsauth}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Forecasting the old-age dependency ratio to determine a sustainable pension age},
  pdfauthor={Rob J Hyndman\^{}\{1*\}, Yijun Zeng\^{}2, Han Lin Shang\^{}3},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
%% Any special functions or other packages can be loaded here.
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

@@

\usepackage{mathptmx}

\raggedbottom
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{anzsj}

@@

Some may argue that using OADR as a measure of a financial burden for the next generation could be biased, with a large number of retirees able to rely on self-financed retirement income, such as superannuation. However, approximately two-thirds of Australians over pension age rely fully or partly on the age pension \citep{AIHW19}, and this figure has remained relatively constant over the last 20 years. Therefore, the proportion of people over working age is still a good representation of the financial burden on future working generations. Nevertheless, using OADR as a single measurement of financial burden may overlook some important economic factors, and \citet{HY12} have suggested the use of an adjusted OADR that takes account of some of these factors in analyzing pension finance. We do not consider these here. Our approach is intended to illustrate how population forecasting can be used to explore the pension age issue in a simple context, and it could be extended to take into account other variables if required.

@@

The ageing trend in Australia, along with retirees' dependence on the age pension, means the cost of the age pension is currently around at 2.7\% of Gross Domestic Product. With an ageing population, the pension age will need to increase or the fiscal burden will become unsustainable.

@@

Following the general procedure proposed in \citet{HB08}, we model each of the five components of population change: female mortality rates, male mortality rates, female net migration, male net migration and fertility rates, and simulate these five components into the future. We use the HU method applied to log-fertility, and the product-ratio method applied to log-mortality and net-migration. We used weighted penalized splines \citep{HU07} when estimating the log-fertility and log-mortality curves, where the weights account for Poisson variation in deaths and births. We use loess when estimating the net-migration curves.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Set \(a_{T+h}=a_{T+h-1}\).
\item
  Increment \(a_{T+h}\) by one-month intervals until either \(\widehat{O}_{T+h|T} \le O^*\) or \(a_{T+h} - a_{T+h-1}=12\) months, where \(\widehat{O}_{T+h|T}\) denotes the mean of the simulated \(O_{T+h|T}\) values.
\end{enumerate}

We repeat these steps for \(h=2,\dots,H\). The second limit in Step 2 arises because the pension age should not jump more than 12 months between years, to prevent pension-age people losing their pension.

### 2021-03-15 PensionAge_wp.tex
Revision submitted

Some may argue that using OADR as a measure of a financial burden for the next generation could be biased, with a large number of retirees able to rely on self-financed retirement income, such as superannuation. However, 80\% of Australians over pension age rely fully or partly on the age pension \autocite{Power14}. Therefore, the proportion of people over working age is still a good representation of the financial burden on future working generations. Nevertheless, using OADR as a single measurement of financial burden may overlook some important economic factors, and \textcite{HY12} have suggested the use of an adjusted OADR that takes account of some of these factors in analyzing pension finance. We do not consider these here. Our approach is intended to illustrate how population forecasting can be used to explore the pension age issue in a simple context, and it could be extended to take into account other variables if required.

@@

Following the general procedure proposed in \textcite{HB08}, we model each of the five components of population change: female mortality rates, male mortality rates, female net migration, male net migration and fertility rates, and simulate these five components into the future. We use the HU method applied to log-fertility, and the product-ratio method applied to log-mortality and net-migration.

>>>>

Some may argue that using OADR as a measure of a financial burden for the next generation could be biased, with a large number of retirees able to rely on self-financed retirement income, such as superannuation. However, approximately two-thirds of Australians over pension age rely fully or partly on the age pension \autocite{AIHW19}, and this figure has remained relatively constant over the last 20 years. Therefore, the proportion of people over working age is still a good representation of the financial burden on future working generations. Nevertheless, using OADR as a single measurement of financial burden may overlook some important economic factors, and \textcite{HY12} have suggested the use of an adjusted OADR that takes account of some of these factors in analyzing pension finance. We do not consider these here. Our approach is intended to illustrate how population forecasting can be used to explore the pension age issue in a simple context, and it could be extended to take into account other variables if required.

@@

Following the general procedure proposed in \textcite{HB08}, we model each of the five components of population change: female mortality rates, male mortality rates, female net migration, male net migration and fertility rates, and simulate these five components into the future. We use the HU method applied to log-fertility, and the product-ratio method applied to log-mortality and net-migration. We used weighted penalized splines \autocite{HU07} when estimating the log-fertility and log-mortality curves, where the weights account for Poisson variation in deaths and births. We use loess when estimating the net-migration curves.
